---
title: "final_project_part1"
output: html_document
date: "2025-11-15"
---
**Name: Miguel Zarate**

Null hypothesis: Crime incidents are distributed normally across a 24 hour time frame. There is no change based on the hour.

Alternative hypothesis: Crime incidents are not evenly distributed across 24 hours. Some times of the day have higher rates of crime than others. 

**What does it contain and what time period does it cover?**
  The dataset I'm analyzing contains crime statistics across Maryland. The time period is 2023-2024.

**Which columns will be most important for your analysis?**
  So far I have looked at incident_hour and offense category name These have the most number that I can use. I have also looked at county_name since a location could gleam information.

**What's missing that you might need or want?**
  Is there a reason why so many crimes happen at midnight? Is it just a tracking error where if a crime has no time associated with it, it just gets assigned to midnight? Is 00:00 just a placeholder for certain crimes?

**Select a relevant column and produce basic descriptive statistics on the mean, median, lowest and highest values.**
  See above.
  

**What questions do you have about it or need help answering?**
  Why does so much crime happen at midnight? 
  Do these crimes change depending on the month?
  Have crime rates been going down?
  
  Other notes: If 00:00 is meant to be a placeholder, how can police ensure more accurate statistics reporting?


```{r}
library(tidyverse)
library(janitor)
```
Creating tables:
```{r}
md_crime <- read_csv("md_crime2023-2024.csv")
population <- read_csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/md_county_pop.csv")

colnames(md_crime)
```

```{r}
summary <- data.frame(
  mean = mean(md_crime$incident_hour, na.rm = TRUE),
  median = median(md_crime$incident_hour, na.rm = TRUE),
  sd = sd(md_crime$incident_hour, na.rm = TRUE)
)
```
Create graphs to visualize times:
```{r}
md_crime %>%
  ggplot(aes(x = incident_hour)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of incident times.",
    x = "Hour of the Day (0–23)",
    y = "Number of Incidents"
  )
```
The visualization shows there is a large amount of crimes listed at 0 hours (midnight). There's also a spike at noon as well, but the midnight spike is much more drastic.

Seeing which crimes are put on the midnight crime spot:
```{r}
md_crime %>%
  filter(incident_hour == 0) %>%
  group_by(offense_category_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Seeing which areas have high midnight crime rates:
```{r}
md_crime %>%
  filter(incident_hour == 0) %>%
  group_by(county_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```
Anova test: do certain crimes occur over different times of day?
```{r}
anova_model <- aov(incident_hour ~ offense_category_name, data = md_crime)
summary(anova_model)
```
```{r}
TukeyHSD(anova_model)
```
```{r}
tuk <- TukeyHSD(aov(incident_hour ~ offense_category_name, data = md_crime))

tuk_df <- as.data.frame(tuk$offense_category_name)

tuk_df$comparison <- rownames(tuk$offense_category_name)

write.csv(tuk_df, "tukey_offense_category.csv", row.names = FALSE)

```

Answer: yes, certain crimes occur much later than others. I think some of the less common crimes are skewing some results. 


Z-score test: Which hours have unusually high crime volumes. 
```{r}
hour_counts <- md_crime %>%
  count(incident_hour)

hour_counts <- hour_counts %>%
  mutate(z = (n - mean(n)) / sd(n))
```
Answer: Midnight is 4 standard deviations from the mean. 
