---
title: "final_project_part1"
output: html_document
date: "2025-11-15"
---
**Name: Miguel Zarate**
**Story pitch: main findings, potential problems, and provide a reporting plan** 
From my research, the main thing I found was that there was a massive spike at midnight for these crimes. The main problem is that midnight could just be a placeholder time. I have two plans for two scenarios. If there is a genuine spike in crimes at midnight, it would be a good idea to find out the type of those crimes that occur at midnight. 30% are larceny, which may be significant. The other story if midnight is a placeholder for non-specific times, we could report on how many crimes don't have a specific time. There's a huge amount of crimes listed at midnight and it modifies the rest of the data since they don't have specific times. This creates an issue for people who study the data. It also show's how non-specific police recordings are, raising that it's important for data to be accurate on crime reports. One solution we could bring up in the article is if the police recorded the minute as well as the hour, which would allow placeholders to be put at 00:00 exactly.

Null hypothesis: Crime incidents are distributed normally across a 24 hour time frame. There is no change based on the hour.

Alternative hypothesis: Crime incidents are not evenly distributed across 24 hours. Some times of the day have higher rates of crime than others. 

**What does it contain and what time period does it cover?**
  The dataset I'm analyzing contains crime statistics across Maryland. The time period is 2023-2024.

**Which columns will be most important for your analysis?**
  So far I have looked at incident_hour and offense category name These have the most number that I can use. I have also looked at county_name since a location could gleam information.

**What's missing that you might need or want?**
  Is there a reason why so many crimes happen at midnight? Is it just a tracking error where if a crime has no time associated with it, it just gets assigned to midnight? Is 00:00 just a placeholder for certain crimes?

**Select a relevant column and produce basic descriptive statistics on the mean, median, lowest and highest values.**
  See above.
  

**What questions do you have about it or need help answering?**
  Why does so much crime happen at midnight? 
  Do these crimes change depending on the month?
  Have crime rates been going down?
  
  Other notes: If 00:00 is meant to be a placeholder, how can police ensure more accurate statistics reporting?


```{r}
library(tidyverse)
library(janitor)
```
Creating tables:
```{r}
md_crime <- read_csv("md_crime2023-2024.csv")
population <- read_csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/md_county_pop.csv")

colnames(md_crime)
```

```{r}
summary <- data.frame(
  mean = mean(md_crime$incident_hour, na.rm = TRUE),
  median = median(md_crime$incident_hour, na.rm = TRUE),
  sd = sd(md_crime$incident_hour, na.rm = TRUE),
  max = max(md_crime$incident_hour, na.rm = TRUE),
  min = min(md_crime$incident_hour, na.rm = TRUE)
)
```
Create graphs to visualize times:
```{r}
md_crime %>%
  ggplot(aes(x = incident_hour)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of incident times.",
    x = "Hour of the Day (0–23)",
    y = "Number of Incidents"
  )
```
The visualization shows there is a large amount of crimes listed at 0 hours (midnight). There's also a spike at noon as well, but the midnight spike is much more drastic. 

Seeing which crimes are put on the midnight crime spot:
```{r}
md_crime %>%
  filter(incident_hour == 0) %>%
  group_by(offense_category_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Seeing which areas have high midnight crime rates:
```{r}
md_crime %>%
  filter(incident_hour == 0) %>%
  group_by(county_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```
Anova test: do certain crimes occur over different times of day? In other words, does average incident hour differ across different times of day?
```{r}
anova_model <- aov(incident_hour ~ offense_category_name, data = md_crime)
summary(anova_model)
```
F-statistic explanation: The f-stat compares variation in incident times between and within incident categories. 
```{r}
TukeyHSD(anova_model)
```
```{r}
tuk <- TukeyHSD(aov(incident_hour ~ offense_category_name, data = md_crime))

tuk_df <- as.data.frame(tuk$offense_category_name)

tuk_df$comparison <- rownames(tuk$offense_category_name)

write.csv(tuk_df, "tukey_offense_category.csv", row.names = FALSE)

```

Answer: yes, certain crimes occur much later than others. I think some of the less common crimes are skewing some results. This confirms the alternative hypothesis and allows us to reject the null hypothesis. The tukey test shows that certain crimes are more likely to happen during certain times based on the data.


Z-score test: Which hours have unusually high crime volumes. 
```{r}
hour_counts <- md_crime %>%
  count(incident_hour)

hour_counts <- hour_counts %>%
  mutate(z = (n - mean(n)) / sd(n))
```
Answer: Midnight is 4 standard deviations from the mean. This confirms the alternative hypothesis.

Listing midnight crimes by type.
```{r}
midnight_crime_percent <- md_crime %>%
  filter(incident_hour == 0) %>%
  count(offense_category_name) %>%
  mutate(percent = round(n / sum(n) * 100, 1)) %>%
  arrange(desc(n))

midnight_crime_percent

```

Finding: Most crimes listed at midnight are theft/larceny.

Creating a pie chart to visualize.
```{r}
midnight_crime <- md_crime %>%
  filter(incident_hour == 0) %>%
  count(offense_category_name, sort = TRUE)

ggplot(midnight_crime, aes(x = "", y = n, fill = offense_category_name)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  labs(
    title = "Distribution of Crimes Recorded at Midnight by Offense Type",
    fill = "Offense Category"
  ) +
  theme_void()
```
